---
title: "FeedForward NN"
output: html_notebook
---

#1

# Brief

In this session I am going to introduce the usage of package 'neuralnet' and associated packages.

Lot's of R packages are built and maintained by a group of users. Since mostly written in R, performance of these packages is sometimes inferior to their industrial alternatives. However, the advantage of using pure-r packages is the ease of use, consistency in computing (if you think of parallel computing implementations), and availability of source code. You can promote the development iof certain R packages by collaborating.

Usually, packages in R focus on certain professional problem and are very specialized. You may get more implemented solutions in Python but the in-depth considerations for certain models are usually better developed in R. There are packages that go really deep into one certain model in some professional aspects--the prediction and representation of which is beneficial. If one want to know as much as a ML model knows, then better representations of such are needed.

There is a misunderstanding that Neural-nets are suitable for all cases.

 - There are different types of NNs and selecting and tuning them takes time
 - Bagged models, boosted models can predict situations with equivalent or even higher complexity sometimes.
 - Complexity of the models raises ambiguity.
 

https://github.com/bips-hb/neuralnet

Another package called 'nnet'.
https://github.com/cran/nnet

For plotting it is useful to call some model-specific visualizer
https://github.com/fawda123/NeuralNetTools

And, as always, visualization for regression models with visreg
https://github.com/pbreheny/visreg



Besides, you can also connect some very powerful platforms, e.g. tf, mxnet to R, but in the case of understanding modernized stats methods, they are not helpful.

 
```{r loadPkg}
library(neuralnet)


```
 
## Train workflow

Let us first agree that a function that can train NN contains (roughly):

 - A Data structure, maybe class,to store the model data
 - A compute function, in this case feed forward, to compute outcome
 - An optimizer (possibly with AD) to find best set of params
 - A function to tune the optimizer
 
 
```{r}
Species_model = neuralnet::neuralnet(Species~., data=iris)
```

Plot structure of the model

```{r}
plot(Species_model)
```

Plot mappings of the model

```{r}
#install.packages("nnet")
library(nnet)
?nnet
Species_model = nnet(Species ~ ., data = iris, subset = samp, size = 2, rang = 0.1,
               decay = 5e-4, maxit = 200)

```

